---
title: "5. A predictive modelling case study"
author: "James"
date: "05/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Tutorial link: https://www.tidymodels.org/start/case-study/

## Introduction

In this final case study, we will use all of the previous articles as a foundation to build a predictive model from beginning to end with data on hotel stays.

```{r}
library(tidymodels)  

# Helper packages
library(readr)       # for importing data
library(vip)         # for variable importance plots
```

## The hotel bookings data

Let’s use hotel bookings data from Antonio, Almeida, and Nunes (2019) to predict which hotel stays included children and/or babies, based on the other characteristics of the stays such as which hotel the guests stay at, how much they pay, etc.

To start, let’s read our hotel data into R, which we’ll do by providing readr::read_csv() with a url where our CSV data is located.

```{r}
library(tidymodels)
library(readr)

hotels <- 
  read_csv('https://tidymodels.org/start/case-study/hotels.csv') %>%
  mutate_if(is.character, as.factor) 

dim(hotels)
#> [1] 50000    23
```

```{r}
glimpse(hotels)
```

We will build a model to predict which actual hotel stays included children and/or babies, and which did not. Our outcome variable `children` is a factor variable with two levels:

```{r}
hotels %>% 
  count(children) %>% 
  mutate(prop = n/sum(n))
```

We can see that children were only in 8.1% of the reservations. This type of class imbalance can often wreak havoc on an analysis. While there are several methods for combating this issue using recipes (search for steps to `upsample` or `downsample`) or other more specialized packages like themis, the analyses shown below analyze the data as-is.

## Data splitting & resampling

For a data splitting strategy, let’s reserve 25% of the stays to the test set. As in our *Evaluate your model with resampling* article, we know our outcome variable `children` is pretty imbalanced so we’ll use a stratified random sample:

```{r}
set.seed(123)
splits      <- initial_split(hotels, strata = children)

hotel_other <- training(splits)
hotel_test  <- testing(splits)

# training set proportions by children
hotel_other %>% 
  count(children) %>% 
  mutate(prop = n/sum(n))

```

```{r}
# test set proportions by children
hotel_test  %>% 
  count(children) %>% 
  mutate(prop = n/sum(n))
```

In our articles so far, we’ve relied on 10-fold cross-validation as the primary resampling method using `rsample::vfold_cv()`. This has created 10 different resamples of the training set (which we further split into analysis and assessment sets), producing 10 different performance metrics that we then aggregated.

For this case study, rather than using multiple iterations of resampling, let’s create a single resample called a validation set. In tidymodels, a validation set is treated as a single iteration of resampling. This will be a split from the 37,500 stays that were not used for testing, which we called `hotel_other`. This split creates two new datasets:

* the set held out for the purpose of measuring performance, called the validation set, and

* the remaining data used to fit the model, called the training set.

We’ll use the `validation_split()` function to allocate 20% of the `hotel_other` stays to the validation set and 30,000 stays to the training set. This means that our model performance metrics will be computed on a single set of 7,500 hotel stays. This is fairly large, so the amount of data should provide enough precision to be a reliable indicator for how well each model predicts the outcome with a single iteration of resampling.

```{r}
set.seed(234)
val_set <- validation_split(hotel_other, 
                            strata = children, 
                            prop = 0.80)
val_set
```

This function, like initial_split(), has the same strata argument, which uses stratified sampling to create the resample. This means that we’ll have roughly the same proportions of hotel stays with and without children in our new validation and training sets, as compared to the original `hotel_other` proportions.

## A first model: Penalised logistic regression

Since our outcome variable children is categorical, logistic regression would be a good first model to start. Let’s use a model that can perform feature selection during training. The glmnet R package fits a generalized linear model via penalized maximum likelihood. This method of estimating the logistic regression slope parameters uses a penalty on the process so that less relevant predictors are driven towards a value of zero. One of the glmnet penalization methods, called the lasso method, can actually set the predictor slopes to zero if a large enough penalty is used.

### Build the model

To specify a penalized logistic regression model that uses a feature selection penalty, let’s use the parsnip package with the glmnet engine:

```{r}
lr_mod <- 
  logistic_reg(penalty = tune(), mixture = 1) %>% 
  set_engine("glmnet")
```

We’ll set the `penalty` argument to `tune()` as a placeholder for now. This is a model hyperparameter that we will tune to find the best value for making predictions with our data. 
Setting `mixture` to a value of one means that the glmnet model will potentially remove irrelevant predictors and choose a simpler model.

### Create the recipe

Let’s create a recipe to define the preprocessing steps we need to prepare our hotel stays data for this model. It might make sense to create a set of date-based predictors that reflect important components related to the arrival date. We have already introduced a number of useful recipe steps for creating features from dates:

* `step_date()` creates predictors for the year, month, and day of the week.

* `step_holiday()` generates a set of indicator variables for specific holidays. Although we don’t know where these two hotels are located, we do know that the countries for origin for most stays are based in Europe.

* `step_rm()` removes variables; here we’ll use it to remove the original date variable since we no longer want it in the model.

Additionally, all categorical predictors (e.g., `distribution_channel`, `hotel`, …) should be converted to dummy variables, and all numeric predictors need to be centered and scaled.

* `step_dummy()` converts characters or factors (i.e., nominal variables) into one or more numeric binary model terms for the levels of the original data.

* `step_zv()` removes indicator variables that only contain a single unique value (e.g. all zeros). This is important because, for penalized models, the predictors should be centered and scaled.

* `step_normalize()` centers and scales numeric variables.

Putting all these steps together into a recipe for a penalized logistic regression model, we have:

```{r}
holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", 
              "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")

lr_recipe <- 
  recipe(children ~ ., data = hotel_other) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = holidays) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())

```


### Create the workflow

As we introduced in *Preprocess your data with recipes*, let’s bundle the model and recipe into a single `workflow()` object to make management of the R objects easier:

```{r}
lr_workflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(lr_recipe)
```

### Create the grid for tuning

Before we fit this model, we need to set up a grid of `penalty` values to tune. In our *Tune model parameters* article, we used `dials::grid_regular()` to create an expanded grid based on a combination of two hyperparameters. Since we have only one hyperparameter to tune here, we can set the grid up manually using a one-column tibble with 30 candidate values:

```{r}
lr_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))

lr_reg_grid %>% top_n(-5) # lowest penalty values

lr_reg_grid %>% top_n(5)  # highest penalty values
```

### Train and tune the model

Let’s use `tune::tune_grid()` to train these 30 penalized logistic regression models. We’ll also save the validation set predictions (via the call to `control_grid()`) so that diagnostic information can be available after the model fit. The area under the ROC curve will be used to quantify how well the model performs across a continuum of event thresholds (recall that the event rate—the proportion of stays including children— is very low for these data).

```{r}
lr_res <- 
  lr_workflow %>% 
  tune_grid(val_set,
            grid = lr_reg_grid,
            control = control_grid(save_pred = TRUE),  # save the val set predictions
            metrics = metric_set(roc_auc))

```

It might be easier to visualize the validation set metrics by plotting the area under the ROC curve against the range of penalty values:

```{r}
lr_plot <- 
  lr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number())

lr_plot 

```

This plots shows us that model performance is generally better at the smaller penalty values. This suggests that the majority of the predictors are important to the model. We also see a steep drop in the area under the ROC curve towards the highest penalty values. This happens because a large enough penalty will remove all predictors from the model, and not surprisingly predictive accuracy plummets with no predictors in the model (recall that an ROC AUC value of 0.50 means that the model does no better than chance at predicting the correct class).

Our model performance seems to plateau at the smaller penalty values, so going by the `roc_auc` metric alone could lead us to multiple options for the “best” value for this hyperparameter:

```{r}
top_models <-
  lr_res %>% 
  show_best("roc_auc", n = 15) %>% 
  arrange(penalty) 

top_models
```

Every candidate model in this tibble likely includes more predictor variables than the model in the row below it. If we used `select_best()`, it would return candidate model 8 with a penalty value of 0.00053, shown with the dotted line below.

However, we may want to choose a penalty value further along the x-axis, closer to where we start to see the decline in model performance. For example, candidate model 12 with a penalty value of 0.00137 has effectively the same performance as the numerically best model, but might eliminate more predictors. In general, fewer irrelevant predictors is better. **If performance is about the same, we’d prefer to choose a higher penalty value.**

Let’s select this value and visualize the validation set ROC curve:

```{r}
lr_best <- 
  lr_res %>% 
  collect_metrics() %>% 
  arrange(penalty) %>% 
  slice(12)

lr_best
```

```{r}
lr_auc <- 
  lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(children, .pred_children) %>% 
  mutate(model = "Logistic Regression")

autoplot(lr_auc)
```

The level of performance generated by this logistic regression model is good, but not groundbreaking. Perhaps the linear nature of the prediction equation is too limiting for this data set. As a next step, we might consider a highly non-linear model generated using a tree-based ensemble method.

### A second model: tree-based ensemble




